---
title: "i-see-your-true-colors-shining-through"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Benjamin Motz"
date: "5/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

For months, people have been naming colors on the internet.  A collaborative effort to create descriptive labels for every color in the RGB/web space exists at <https://colornames.org>, where visitors can propose new names for colors, upvote or downvote existing names, or download the full set of all named colors.  When a user names a color, they are invited to name a new color, selected at random.  There is no instruction other than to "Please be descriptive and creative."  

The names provided offer a spectacular wealth of information about psychological associations across the color spectrum.  For example, are there patterns in the sentiment of color names for different hues?  Do people associate different colors with different emotions? 

On May 2, 2020, Guy K. from <https://colornames.org> generously provided a dump of the complete colornames.org database, enhanced with timestamps of when names were proposed and votes were cast, and including geolocation data (inferred from visitor IP addresses using the MaxMind GeoIP2 lookup table).  This markdown file documents analyses of these data.

## Initial Import and Preprocessing

We'll start with some basic cleanup and initiating libraries.

```{r frontmatter, results='hide', message = FALSE, warning = FALSE}
# Clean up
graphics.off()
rm(list=ls())
set.seed(100)
# Libraries
require(tidyverse)  # Keepin' it real
require(fields)     # Creating smooth surfaces
require(reshape2)   # Melting surface matrix back into tidy format
require(lutz)       # Lookup time zone from lat/lon point
require(lubridate)  # Easy date manipulation
require(sentimentr) # Estimate sentiment of text
```

Now let's download the copy of the raw colornames.org data from OSF:

```{r colornames-download}
temp <- tempfile()
download.file("https://osf.io/p25f9/download",temp) # variants - all name proposals
data <- read.csv(temp,header=FALSE,col.names=c("hexcode", "name", "votes", "submit_date", "user_id", "country_iso", "subdivision_iso", "city_name", "postal_code", "latitude", "longitude", "accuracy_radius"),stringsAsFactors=FALSE)
```

Fix things, make things as they should be:

```{r preprocessing, results='hide', message = FALSE, warning = FALSE}
# Remove proposals by the publisher, which were only intended to seed the initiative
data <- data %>% filter(user_id != "d7de10d8a0ee2a852de36b128ef7eaac9a39f6f6") 
# Remove errant values (maybe blank proposals?)
data <- data %>% filter(name != "")
# Add RGB and HSV
data <- cbind(data,t(col2rgb(paste("#",data$hexcode,sep = "")))) # RGB
data <- cbind(data,t(rgb2hsv(data$red,data$green,data$blue))) # HSV
# Fix the latitude and longitude
data$latitude <- as.numeric(data$latitude)
data$longitude <- as.numeric(data$longitude)
data$latitude[data$city_name=="Comandante Nicanor Otamendi"] <- -38.1167
data$longitude[data$city_name=="Comandante Nicanor Otamendi"] <- -57.85
# Add timezone to the data
data <- data %>% mutate(tz = tz_lookup_coords(latitude, longitude, method = "fast"))
# Convert submission timestamp strings into timestamps
data$submit_date <- ymd_hms(data$submit_date)
# For those records where we have a time zone, force it on a new date column
local.data <- data %>% filter(!is.na(tz)) %>% rowwise %>% 
  mutate(local_submit_date = force_tz(submit_date,tz)) %>%
  select(hexcode,name,votes,user_id,submit_date,local_submit_date)
data <- merge(data, local.data, by=c("hexcode","name","votes","user_id","submit_date"), all.x = TRUE)
# Remove duplicates
data <- data %>% distinct(hexcode,name,votes,user_id,submit_date,local_submit_date, .keep_all = TRUE)
```

## Descriptive Analyses

With the data all imported and preprocessed, we can do some basic summaries of the dataset, just to better-understand what we're dealing with.

```{r descriptive-analyses}
# How many colorname proposals over time?
data %>% mutate(date = as.Date(submit_date,"GMT")) %>% 
  ggplot(aes(x=date)) + 
  geom_histogram(binwidth=30)
# What time of day did these get submitted?
data %>% filter(!is.na(local_submit_date)) %>% mutate(hour = hour(local_submit_date)) %>%
  ggplot(aes(x=hour)) +
  geom_histogram(binwidth=1)
# Where are these proposals coming from?
ggplot() + borders("world", colour="gray50", fill="gray50") +
  geom_point(aes(x=data$longitude,y=data$latitude), color="blue", size=1)

```

## Sentiment

The colornames provided can be estimated for their valence (positive/negative) of sentiment <https://en.wikipedia.org/wiki/Sentiment_analysis>.  There is a well-established body of work examining color preferences across the color spectrum.  The sentiment of color names may provide complementary insights into humans' preferences for different colors.

```{r sentiment-analyses}
# Get sentiment from sentimentr; this takes a long while
data <- data %>% select(name) %>% get_sentences() %>% 
  sentiment_by() %>% select(sentiment = ave_sentiment) %>% bind_cols(data,.)
# Sentiment over time of day
# Noteworthy that there's nothing here really.  
data %>% filter(!is.na(local_submit_date)) %>% mutate(hour = hour(local_submit_date)) %>%
  group_by(hour) %>% summarize(avg_sentiment = mean(sentiment)) %>%
  ggplot(aes(x=hour, y=avg_sentiment)) + 
  geom_line()
# Recreate Fig 1C from Palmer & Schloss (2010): 10.1073/pnas.0906172107
# Define the centroid of the 32 colors in the Berkeley Color Project
# "S" is saturated, "L" is light, "M" is muted, "D" is dark
# "R"ed "O"range "Y"ellow c"H"artreuse "G"reen "C"yan "B"lue "P"urple
palmerCats <- data.frame(slice=character(32),hue=character(32),
                         r=numeric(32),g=numeric(32),b=numeric(32),
                         stringsAsFactors = FALSE)
palmerCats[1, ] <- list("Saturated","1",238,0,89)
palmerCats[2, ] <- list("Saturated","2",246,146,36)
palmerCats[3, ] <- list("Saturated","3",255,237,14)
palmerCats[4, ] <- list("Saturated","4",182,218,56)
palmerCats[5, ] <- list("Saturated","5",105,200,133)
palmerCats[6, ] <- list("Saturated","6",96,202,213)
palmerCats[7, ] <- list("Saturated","7",104,164,222)
palmerCats[8, ] <- list("Saturated","8",157,69,159)
palmerCats[9, ] <- list("Light","1",246,145,188)
palmerCats[10,] <- list("Light","2",255,204,168)
palmerCats[11,] <- list("Light","3",255,240,158)
palmerCats[12,] <- list("Light","4",228,238,151)
palmerCats[13,] <- list("Light","5",119,232,200)
palmerCats[14,] <- list("Light","6",167,224,233)
palmerCats[15,] <- list("Light","7",175,198,236)
palmerCats[16,] <- list("Light","8",184,158,201)
palmerCats[17,] <- list("Muted","1",207,115,144)
palmerCats[18,] <- list("Muted","2",211,155,119)
palmerCats[19,] <- list("Muted","3",223,205,116)
palmerCats[20,] <- list("Muted","4",182,207,99)
palmerCats[21,] <- list("Muted","5",134,208,147)
palmerCats[22,] <- list("Muted","6",139,211,214)
palmerCats[23,] <- list("Muted","7",130,160,207)
palmerCats[24,] <- list("Muted","8",164,112,170)
palmerCats[25,] <- list("Dark","1",164,21,68)
palmerCats[26,] <- list("Dark","2",160,89,47)
palmerCats[27,] <- list("Dark","3",161,154,55)
palmerCats[28,] <- list("Dark","4",129,158,68)
palmerCats[29,] <- list("Dark","5",45,158,113)
palmerCats[30,] <- list("Dark","6",42,159,159)
palmerCats[31,] <- list("Dark","7",66,126,187)
palmerCats[32,] <- list("Dark","8",118,48,150)
distThresh <- 30
fdat <- data %>% mutate(sbin=NA,hbin=NA)
for (i in 1:nrow(palmerCats)) { # Cheap way of binning
  fdat$sbin[between(fdat$red, palmerCats$r[i]-distThresh, palmerCats$r[i]+distThresh) &
              between(fdat$green, palmerCats$g[i]-distThresh, palmerCats$g[i]+distThresh) &
              between(fdat$blue, palmerCats$b[i]-distThresh, palmerCats$b[i]+distThresh)] <- palmerCats$slice[i]
  fdat$hbin[between(fdat$red, palmerCats$r[i]-distThresh, palmerCats$r[i]+distThresh) &
              between(fdat$green, palmerCats$g[i]-distThresh, palmerCats$g[i]+distThresh) &
              between(fdat$blue, palmerCats$b[i]-distThresh, palmerCats$b[i]+distThresh)] <- palmerCats$hue[i]
}
cat_data <- fdat %>% filter(!is.na(sbin)) %>% group_by(sbin,hbin) %>% 
  summarize(red = mean(red),
            green = mean(green),
            blue = mean(blue),
            sentiment = mean(sentiment))
cat_data$sbin <- factor(cat_data$sbin, levels=c("Saturated","Light","Muted","Dark"))
cat_data$hex <- rgb(round(cat_data$red,0), round(cat_data$green,0), round(cat_data$blue,0), maxColorValue = 255)
ggplot(cat_data,aes(x=hbin,y=sentiment,shape=sbin,group=1)) +
  geom_point(color=cat_data$hex,size=3) +
  scale_shape_manual(values=c(16, 17, 18,15)) + 
  geom_path(aes(group=sbin),size=0.5) +
  scale_x_discrete(labels=c("1"="R","2"="O","3"="Y","4"="H","5"="G","6"="C","7"="B","8"="P")) +
  xlab("Hue") + ylab("Sentiment") + labs(shape="")
```

Make the full surface map of all sentiments for all colors...

```{r sentiment-surface}
m <- data %>% mutate(hbin=ntile(h,11),
                     sbin=ntile(s,11)) %>% 
  group_by(hbin,sbin) %>% summarize(sentiment=mean(sentiment)) %>% 
  pivot_wider(names_from = sbin, values_from = sentiment) %>% ungroup() %>% select(-hbin) %>% as.matrix()
obj <- list(x=0:10/10,y=0:10/10,z=m)
surf <- interp.surface.grid(obj=obj,grid.list=list(x=seq(0,1,length=101),
                                                   y=seq(0,1,length=101)))
dimnames(surf$z) <- list(x=surf$x,y=surf$y)
meltedsurf <- melt(surf,value.name="z") %>% na.omit() %>% select(-L1)
ggplot(meltedsurf,aes(x=x,y=y,z=z,fill=z)) +
  geom_tile() + 
  scale_fill_gradient2(name="Sentiment") +
  xlab("Hue") + ylab("Saturation")  
```